---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

```{r}
# The Prince George's County crime data reveals many different categories, but some crimes are a lot more prevalent each month than others. From a reporting standpoint, a category that would be more important than others would include homicide, sex offense, or any of the assault categories. While reporting all crime is important, those categories heavily impact public safety. Additionally, the legal action taken for these crimes could be a complex story, leading to many follow up articles if necessary. A category that would make the job of summarizing the data more difficult would be a category that includes theft. This could make the job of summarizing data more difficult over others because a financial crime may include subsequent information and documents, which may be difficult to obtain.
```


### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |> mutate(total_incidents = Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec, average_incidents = total_incidents/12)
                                     
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

```{r}
#The categories that have the highest average number of incidents per month include accident, accident with impound, auto stolen, and theft from auto. The categories that have the lowest average number of incidents per month include assault shooting and weapon, B & E other, homicide, and sex offense. This indicates a pattern of vehicle crime in Prince George's County, as well as property crime. High level crimes, such as homicide, have a lower average number of incidents. Thus, a question I would have about a category that has a high average number of incidents per month, such as the accident category is: What is a common reason for accidents in Prince George's County? Could a high traffic volume from a high number of commuters in Prince George's County be a reason?
```


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(`june_change` = (Jun-average_incidents)/average_incidents * 100, `july_change` = (Jul-average_incidents)/average_incidents * 100, `august_change` = (Aug-average_incidents)/average_incidents * 100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?


```{r}
# A pattern I observed for the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August include many negative percentages. While not every category saw a negative percent change, the one's that did indicate a decrease in crime. Categories including B & E other and theft decreased in percentage for all three of these months. This tells me that the theory that crime is more likely to occur in the summer months is not accurate.
```

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

```{r}

# Driven by high traffic volume, Prince George's County, Maryland, reported 5,776 accidents in 2024, averaging about 481 accidents per month.

# Local residents who commute for work or school increase the traffic density in the county during rush hour. More cars forced together on busy highways results in stressful conditions on the road and potential traffic violations. According to the Prince George's County crime data, peak accident numbers were reported in May, October and December.
```


### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
